---
title: "GDP"
execute:
  warning: false
  message: false
---

```{r setup}
# ── Source all required R modules ─────────────────────────────────────────────
# 00_config.R    → shared package loads (dplyr, ggplot2, etc.)
# 01_helpers.R   → caching and environment variable utilities
# API_bea.R      → raw BEA API connector
# mock_data.R    → synthetic data used when USE_MOCK = TRUE
# parse_gdp.R    → parse_bea_gdp() + get_gdp_bundle()
# viz_timeseries.R → maybe_interactive() wrapper for plotly
# viz_gdp.R      → plot_gdp_growth(), plot_gdp_forecast()
# forecast_nextq.R → forecast_next_quarter() ARIMA model

source("R/00_config.R")
source("R/01_helpers.R")
source("R/API_bea.R")
source("R/mock_data.R")
source("R/parse_gdp.R")
source("R/viz_timeseries.R")
source("R/viz_gdp.R")
source("R/forecast_nextq.R")

# ── Mock vs. live data ────────────────────────────────────────────────────────
# USE_MOCK auto-detects whether BEA_API_KEY is available.
# Locally: set BEA_API_KEY in ~/.Renviron
# In CI:   set BEA_API_KEY in GitHub Secrets
USE_MOCK    <- Sys.getenv("BEA_API_KEY") == ""
INTERACTIVE <- FALSE   # flip to TRUE for plotly hover/zoom
```

## Quarterly GDP Growth (Full History)

```{r gdp-load}
# Fetch (or mock) real GDP growth — annualized % change from prior quarter.
# Output: standard tibble — date (first month of quarter), value (%), series_id,
#         label, freq = "Q"
gdp <- get_gdp_bundle(
  start_year = 2000,
  use_mock   = USE_MOCK,
  use_cache  = TRUE,
  cache      = TRUE
)
```

```{r plot-gdp-growth}
# Chart 1: Bar chart of annualized GDP growth by quarter.
# Green = expansion, red/pink = contraction.
# COVID collapse (Q2 2020) and rebound (Q3 2020) are clearly visible.
# To use this chart alone in a presentation: copy just this chunk.
p_gdp_growth <- plot_gdp_growth(gdp)
maybe_interactive(p_gdp_growth, INTERACTIVE)
```

## 3-Quarter GDP Forecast

```{r plot-gdp-forecast}
# Chart 2: Recent GDP quarters + ARIMA model forecast for the next 3 quarters.
# history_n = 20 shows the last 5 years (20 quarters) for context.
# Shaded bands represent 80% and 95% confidence intervals.
# To use this chart alone in a presentation: copy just this chunk.
p_gdp_forecast <- plot_gdp_forecast(gdp, history_n = 20)
maybe_interactive(p_gdp_forecast, INTERACTIVE)
```

---

::: {.callout-note}
**Data source:** U.S. Bureau of Economic Analysis (BEA), National Income and
Product Accounts (NIPA) Table 1.1.1 — Percent Change from Preceding Period in
Real Gross Domestic Product (BEA series A191RL). Seasonally Adjusted
Annual Rate (SAAR). Advance estimate released ~4 weeks after quarter-end;
preliminary and final revisions follow in subsequent months. Forecast produced
by an ARIMA model; confidence intervals reflect model uncertainty only and
should not be interpreted as policy projections.
:::
