---
title: "Inflation"
execute:
  warning: false
  message: false
---

```{r setup}
# ── Source all required R modules ─────────────────────────────────────────────
# 00_config.R    → shared package loads (dplyr, ggplot2, etc.)
# 01_helpers.R   → caching and environment variable utilities
# API_bls.R      → raw BLS API connector (POST request, returns JSON)
# mock_data.R    → synthetic data used when USE_MOCK = TRUE
# parse_cpi.R    → cleans BLS JSON → standard tibble; provides get_cpi_bundle()
# viz_timeseries.R → maybe_interactive() wrapper for plotly
# viz_inflation.R → plot_cpi_index(), plot_cpi_yoy(), plot_cpi_forecast()
# forecast_nextq.R → forecast_next_quarter() ARIMA model

source("R/00_config.R")
source("R/01_helpers.R")
source("R/API_bls.R")
source("R/mock_data.R")
source("R/parse_cpi.R")
source("R/viz_timeseries.R")
source("R/viz_inflation.R")
source("R/forecast_nextq.R")

# ── Mock vs. live data ────────────────────────────────────────────────────────
# USE_MOCK auto-detects whether BLS_API_KEY is available.
# Locally: set BLS_API_KEY in ~/.Renviron
# In CI:   set BLS_API_KEY in GitHub Secrets
USE_MOCK    <- Sys.getenv("BLS_API_KEY") == ""
INTERACTIVE <- FALSE   # flip to TRUE for plotly hover/zoom
```

## Price Index Level

```{r cpi-load}
# Fetch (or mock) the CPI-U All Items, Seasonally Adjusted series.
# Output: standard tibble — date, value (index), series_id, label, freq
cpi <- get_cpi_bundle(
  start_year = 2000,
  use_mock   = USE_MOCK,
  use_cache  = TRUE,
  cache      = TRUE
)
```

```{r plot-cpi-index}
# Chart 1: Raw index level — shows cumulative price growth over time.
# To use this chart alone in a presentation: copy just this chunk.
p_cpi_index <- plot_cpi_index(cpi)
maybe_interactive(p_cpi_index, INTERACTIVE)
```

## Year-over-Year Inflation Rate

```{r plot-cpi-yoy}
# Chart 2: YoY % change — the "headline inflation" number reported in the news.
# Dashed line marks the Fed's 2% inflation target.
# To use this chart alone in a presentation: copy just this chunk.
p_cpi_yoy <- plot_cpi_yoy(cpi)
maybe_interactive(p_cpi_yoy, INTERACTIVE)
```

## 3-Month Inflation Forecast

```{r plot-cpi-forecast}
# Chart 3: Recent YoY inflation + ARIMA model forecast for the next 3 months.
# Shaded bands represent 80% and 95% confidence intervals.
# history_n = 60 shows the last 5 years for readability.
# To use this chart alone in a presentation: copy just this chunk.
p_cpi_forecast <- plot_cpi_forecast(cpi, history_n = 60)
maybe_interactive(p_cpi_forecast, INTERACTIVE)
```

---

::: {.callout-note}
**Data source:** U.S. Bureau of Labor Statistics (BLS), CPI-U All Items
Seasonally Adjusted (series CUSR0000SA0). Released monthly, typically in the
second week of the following month. Forecast produced by an ARIMA model fit to
the YoY % change series; confidence intervals reflect model uncertainty only
and do not incorporate external shocks.
:::
