---
title: "Labor Market"
execute:
  warning: false
  message: false
---

```{r setup}
# ── Source all required R modules ─────────────────────────────────────────────
# 00_config.R    → shared package loads (dplyr, ggplot2, etc.)
# 01_helpers.R   → caching and environment variable utilities
# API_bls.R      → raw BLS API connector
# mock_data.R    → synthetic data used when USE_MOCK = TRUE
# parse_cpi.R    → provides parse_bls_series() (shared parser, sourced first)
# parse_labor.R  → get_labor_bundle() wrapper
# viz_timeseries.R → maybe_interactive() wrapper for plotly
# viz_labor.R    → plot_unrate(), plot_unrate_forecast()
# forecast_nextq.R → forecast_next_quarter() ARIMA model

source("R/00_config.R")
source("R/01_helpers.R")
source("R/API_bls.R")
source("R/mock_data.R")
source("R/parse_cpi.R")      # provides parse_bls_series() used by parse_labor.R
source("R/parse_labor.R")
source("R/viz_timeseries.R")
source("R/viz_labor.R")
source("R/forecast_nextq.R")

# ── Mock vs. live data ────────────────────────────────────────────────────────
# USE_MOCK auto-detects whether BLS_API_KEY is available.
# Locally: set BLS_API_KEY in ~/.Renviron
# In CI:   set BLS_API_KEY in GitHub Secrets
USE_MOCK    <- Sys.getenv("BLS_API_KEY") == ""
INTERACTIVE <- FALSE   # flip to TRUE for plotly hover/zoom
```

## Unemployment Rate (Full History)

```{r labor-load}
# Fetch (or mock) the U-3 Unemployment Rate, Seasonally Adjusted.
# Output: standard tibble — date, value (%), series_id, label, freq
labor <- get_labor_bundle(
  start_year = 2000,
  use_mock   = USE_MOCK,
  use_cache  = TRUE,
  cache      = TRUE
)
```

```{r plot-unrate}
# Chart 1: Full unemployment rate history with NBER recession shading.
# Grey bands mark officially designated recession periods.
# To use this chart alone in a presentation: copy just this chunk.
p_unrate <- plot_unrate(labor)
maybe_interactive(p_unrate, INTERACTIVE)
```

## 3-Month Unemployment Forecast

```{r plot-unrate-forecast}
# Chart 2: Recent unemployment + ARIMA model forecast for the next 3 months.
# history_n = 60 shows the last 5 years for readability.
# Shaded bands represent 80% and 95% confidence intervals.
# To use this chart alone in a presentation: copy just this chunk.
p_unrate_forecast <- plot_unrate_forecast(labor, history_n = 60)
maybe_interactive(p_unrate_forecast, INTERACTIVE)
```

---

::: {.callout-note}
**Data source:** U.S. Bureau of Labor Statistics (BLS), Current Population
Survey (CPS). U-3 Unemployment Rate, Seasonally Adjusted (series LNS14000000).
Released monthly on the first Friday of the following month. Recession shading
based on NBER Business Cycle Dating Committee designations. Forecast produced
by an ARIMA model; confidence intervals reflect model uncertainty only.
:::
